---
title: "README"
author: "Ren Yi Wang"
date: "April 27, 2018"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


## Overview

`"dieroller"` is a minimal [R](http://www.r-project.org/) package 
that provides functions to simulate rolling a die.

* `die()` creates a coin object (of class `"die"`)
* `roll()` tosses a coin object, producing a `"roll"` object.
* `plot()` method for a `"roll"` object to plot frequencies of heads.
* `summary()` method for a `"roll"` object.


## Motivation

This package has been developed to illustrate some of the concepts
behind the creation of an R package.


## Installation

Install the development version from GitHub via the package `"devtools"`:

```r
# development version from GitHub:
#install.packages("devtools") 

# install "cointoss" (without vignettes)
devtools::install_github("gastonstat/cointoss")

# install "cointoss" (with vignettes)
devtools::install_github("gastonstat/cointoss", build_vignettes = TRUE)
```


## Usage

```{r}
library(dieroller)

# default die
fair_die <- die()
fair_die

# 1 roll of fair_die
roll(fair_die, 1)

# 10 rolls of fair_die
roll10 <- roll(fair_die, times = 10)
roll10

# summary
summary(roll10)


# 100 rolls
roll100 <- roll(fair_die, times = 100)

# summary
summary(roll100)
```