---
title: "Stat135_lab_1"
author: "Ren Yi Wang"
date: "February 10, 2018"
output: html_document
---
<font size = "5">

## Part 1 a

```{r}
load("/Users/user/Desktop/Stat 135/KaiserBabies.rda")
set.seed(7)

n <- 10
N <- length(na.omit(infants$wt))

mySample <- sample(na.omit(infants$wt), n)
sampleAverage <- mean(mySample)

# Calculate the 95% interval
estimatedSE <- sqrt((1-(n/N))*var(mySample)/n)
lowerBound <- sampleAverage-1.96*estimatedSE
upperBound <- sampleAverage+1.96*estimatedSE
```

My sample average is `r sampleAverage`, and my estimated SE is `r estimatedSE`. The 95% confidence interval is (`r lowerBound`, `r upperBound`), under the assumption that normality works.

## Part 1 b

```{r}
# Sample 1000 times
sampleVec <- matrix(0, 1, 1000)
record <- matrix(0, 2, 1000)
for (i in 1:1000)
{
  repeatedSample <- sample(na.omit(infants$wt), n)
  repSampleAve <- mean(repeatedSample)
  newEstimatedSE <- sqrt((1-(n/N))*var(repeatedSample)/n)
  sampleVec[1, i] <- repSampleAve
  record[1, i] <- repSampleAve-1.96*newEstimatedSE
  record[2, i] <- repSampleAve+1.96*newEstimatedSE
}

# Calculate the true average in population (exclude na)
trueAverage <- mean(na.omit(infants$wt))

# Cound number of invervals covering true average
coverNumber <- 0
for (k in 1:1000)
{
  if (record[1, k] < trueAverage && trueAverage < record[2, k])
    coverNumber <- coverNumber + 1
  else
    coverNumber <- coverNumber
}
```

I expect 950 inervals to cover the true population average, and `r coverNumber` of my confidence intervals cover true popultion average.

## Part 1 c

```{r}
sampleSD <- sd(sampleVec)
hist(sampleVec, breaks = 50)
qqnorm(sampleVec)
qqline(sampleVec, col = "red", lwd = 2)
```

The SD of sample average is `r sampleSD`, and it is not quite similar to my estimated SE in part a. From my qqplot, sample average is close to the red line, and thus it is close to a normal distribution. Since sample average is close to normal distribution, the confidence interval is valid.

## Part 2 a

```{r}
# Bootstrap function
bootStrap = function(mySample, popSize = 1236, B = 1000, replace =
              FALSE)
{
  if (replace) 
    return(replicate(B, mean(sample(mySample, length(mySample),
      TRUE))))
  else 
    vals = sort(unique(mySample))
    counts = table(mySample)
    bootPop = rep(vals, round(counts * popSize / length(mySample)))
    return(list(bootPop, 
      bootSamps = replicate(B,mean(sample(bootPop, 
        length(mySample), FALSE)))))
}

bootSampAvgs <- bootStrap(mySample)$bootSamps
hist(bootSampAvgs, breaks = 50)
abline(v = mean(bootStrap(mySample)[[1]]),col = "red", lwd = 2)
bootSD <- sd(bootSampAvgs)
```

The SD of sample average is `r bootSD`, and it is close to the estimated SE in part 1 a.

## Part 2 b

```{r}
# Contruct 95% confidence interval
bootLowerBound <- unname(quantile(bootSampAvgs, 0.025))
bootUpperBound <- unname(quantile(bootSampAvgs, 0.975))
```

My 95% confidence interval is (`r bootLowerBound`, `r bootUpperBound`), and it is close to the confidence interval I have in part 1 a.

## Part 3.1.a

```{r}
set.seed(7)

n1 <- 100

mySample1 <- sample(na.omit(infants$wt), n1)
sampleAverage1 <- mean(mySample1)

# Calculate the 95% interval
estimatedSE1 <- sqrt((1-(n1/N))*var(mySample1)/n1)
lowerBound1 <- sampleAverage1-1.96*estimatedSE1
upperBound1 <- sampleAverage1+1.96*estimatedSE1
```

My sample average is `r sampleAverage1`, and my estimated SE is `r estimatedSE1`. The 95% confidence interval is (`r lowerBound1`, `r upperBound1`).

## Part 3.1.b

```{r}
# Sample 1000 times
sampleVec1 <- matrix(0, 1, 1000)
record1 <- matrix(0, 2, 1000)
for (i in 1:1000)
{
  repeatedSample1 <- sample(na.omit(infants$wt), n1)
  repSampleAve1 <- mean(repeatedSample1)
  newEstimatedSE1 <- sqrt((1-(n1/N))*var(repeatedSample1)/n1)
  sampleVec1[1, i] <- repSampleAve1
  record1[1, i] <- repSampleAve1-1.96*newEstimatedSE1
  record1[2, i] <- repSampleAve1+1.96*newEstimatedSE1
}

# Cound number of invervals covering true average
coverNumber1 <- 0
for (k in 1:1000)
{
  if (record1[1, k] < trueAverage && trueAverage < record1[2, k])
    coverNumber1 <- coverNumber1 + 1
  else
    coverNumber1 <- coverNumber1
}
```

I expect 950 inervals to cover the true population average. `r coverNumber1` of my confidence intervals cover true popultion average.

## Part 3.1.c

```{r}
sampleSD1 <- sd(sampleVec1)
hist(sampleVec1, breaks = 50)
qqnorm(sampleVec1)
qqline(sampleVec1, col = "red", lwd = 2)
```

The SD of sample average is `r sampleSD1`, and it is quite similar to my estimated SE in part a. From my qqplot, sample average is close to the red line, and thus it is close to a normal distribution. Since sample average is close to normal distribution, the confidence interval is valid.

## Part 3.2.a

```{r}
bootSampAvgs1 <- bootStrap(mySample1)$bootSamps
hist(bootSampAvgs1, breaks = 50)
abline(v = mean(bootStrap(mySample1)[[1]]),col = "red", lwd = 2)
bootSD1 <- sd(bootSampAvgs1)
```

The SD of sample average is `r bootSD1`, and it is close to the estimated SE in part 3.1.a.

## Part 3.2.b

```{r}
# Contruct 95% confidence interval
bootLowerBound1 <- unname(quantile(bootSampAvgs1, 0.025))
bootUpperBound1 <- unname(quantile(bootSampAvgs1, 0.975))
```

My 95% confidence interval is (`r bootLowerBound1`, `r bootUpperBound1`), and it is close to the confidence interval I have in part 3.1.a.

